---
title: "Vinho Verde wine report"
author: "Raquel Parra Suazo & Shenbin Zheng"
date: "March 28, 2019"
output: word_document
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
packages = c("e1071","gridExtra","shinythemes","shinyLP","tidyverse","shiny","DT","mice","BH","ggplot2","scales","devtools","Rcpp","rCharts","markdown","data.table","dplyr","shinyjs","plotly","vembedr","caret","rpart")
lapply(packages, require, character.only = TRUE)
```

## Introduction 

This Shiny App is for searching, visualizating and extracting useful information about the Portuguese Vinho Verde wine variety. Our aim was to create a way for the user to understand this kind of wine and help them choose the correct variant to combine with their food by means of a predictive model (decision tree).

## Data Description

As we said, the dataset is related to red and white variants of the Portuguese "Vinho Verde" wine. The total number of observations is 6497 where 6031 were useful for developing the app. The proportion of white wines is much higher that the red wines one. In fact, within 6031 observations there are 1570 red wines and 4461 white wines.

```{r data_load, include=FALSE}
url="http://halweb.uc3m.es/esp/Personal/personas/imolina/esp/Archivos/VinhoVerdeQuality_Data.csv"

vino=read.csv(url,header=TRUE,sep=";")
vino <- as_tibble(vino)
vino=vino %>%select(-starts_with("X"))
vino=vino %>%  drop_na()
var_vino=vino %>% select(-Taste,-pH,-Variant,-quality)
scale_vino=vino %>% select(-Taste,-Variant)
```

## Data exploration

First of all, we would like to know what variables does the dataset contain.  

```{r exploration, echo=FALSE}
head(vino)
str(vino)
```

From the output above we know that we have 14 variables in total, where 2 variables ("Variant" and "Taste") are factors and the rest are numerical variables. The function **summary()** also provides useful information: 

```{r exploration2, echo=FALSE}
summary(vino)
```

Here we can see additional information from each variable: median, mean, quantile, etc.

## Data representation

We aim to understand better the variables by means of ploting several graphs. We will focus on white wines, since it has more observations, but in the Shiny App the user can select which wine variant they find of interest in a interactive way.

Here are ploted two pie charts that helps us see the distribution of "Taste" and "Quality":

```{r piechar, echo=FALSE}
#Taste distribution
x1=nrow(vino %>% filter(Taste== 'Balanced', Variant=="white")) 
x2=nrow(vino %>% filter(Taste== 'Light-Bodied', Variant=="white")) 
x3=nrow(vino %>% filter(Taste== 'Low acid', Variant=="white"))
x4=nrow(vino %>% filter(Taste== 'Sweet', Variant=="white"))
x5=nrow(vino %>% filter(Taste== 'Very low acid', Variant=="white"))
total.white=nrow(vino %>% filter(Variant=="white"))
value = c(x1,x2,x3,x4,x5)/total.white
      
df <- data.frame(Taste = c("Balanced", "Light-Bodied", "Low acid", 
                           "Sweet", "Very low acid"),
                 value = c(x1,x2,x3,x4,x5)/total.white)
      
p1 = ggplot(df, aes(x = "", y = value, fill = Taste)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  coord_polar("y", start = 0)+
  theme_void()
    
#Quality distribution
x3=nrow(vino %>% filter(quality== 3, Variant=="white"))
x4=nrow(vino %>% filter(quality== 4, Variant=="white"))
x5=nrow(vino %>% filter(quality== 5, Variant=="white"))
x6=nrow(vino %>% filter(quality== 6, Variant=="white"))
x7=nrow(vino %>% filter(quality== 7, Variant=="white"))
x8=nrow(vino %>% filter(quality== 8, Variant=="white"))
x9=nrow(vino %>% filter(quality== 9, Variant=="white"))
total.white=nrow(vino %>% filter(Variant=="white"))
value = c(x3,x4,x5,x6,x7,x8,x9)/total.white
      
df <- data.frame(Quality = c("Quality=3", "Quality=4", "Quality=5", 
                             "Quality=6", "Quality=7","Quality=8","Quality=9"),
                 value = c(x3,x4,x5,x6,x7,x8,x9)/total.white)
      
p2 = ggplot(df, aes(x = "", y = value, fill = Quality)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  coord_polar("y", start = 0)+
  theme_void()
grid.arrange(p1, p2, ncol=2)
```

After that, we check any significant values, such as the mean of each variable. For that we use a boxplot, differentiating between the six tastes that we have. Here we will represent only one boxplot, corresponding to the variable "fixed.acidity":

```{r boxplot, echo=FALSE}
#par(mfrow=c(1,1))
p<-ggplot(vino,aes_string(x="Taste",y=vino$fixed.acidity))+
      geom_boxplot(aes(color=Taste),outlier.shape = NA)+
      theme_minimal()
#p<-ggplotly(p)
p
```

The next plot we have done is somehow a density plot of each variable and differenced according to the 2 wine variants that we have. Again, we will focus only on one variable, which is "fixed.acidity":

```{r density, echo=FALSE, warning=FALSE}
p <- ggplot(vino, aes_string(x = vino$fixed.acidity)) + 
      geom_bar(aes(y = ..count.., fill = Variant)) +
      scale_fill_brewer(palette = "Set2") + 
      ylab("Number of wines") + 
      theme_minimal()
      
#p <- ggplotly(p)
p
```

And a final plot which is a correlation plot between variables and each point is filled with colors according to which taste it belongs. We will represent here the correlation plot in "fixed.acidity", which will be a lineal plot because the correlation coefficient between one variable with itself is 1:

```{r correlation, echo=FALSE, warning=FALSE}
p=ggplot(vino, aes_string(x=vino$fixed.acidity, y=vino$fixed.acidity,
                          color=vino$Taste)) +
  geom_point(size=2, shape=23)+
  geom_smooth(method="lm", se=TRUE, fullrange=TRUE)+
  theme_minimal()
#ggplotly(p)
p
```

As stated before, our final aim is to predict the wine variant (binary prediction) that the user desires according to his/her interests. We have created the model based on the decision tree model and we will check that the performance of this model is very good (with high prediction accuracy). 

```{r prediction, echo=FALSE}
#DT: Decision Tree

#New dataset with the most important variables
data.pred=vino %>% select(c("fixed.acidity", "residual.sugar", "pH",
                              "alcohol", "quality", "Variant", "Taste"))

#Split up the dataset into two parts: one for training and one for testing
spl1 = createDataPartition(data.pred$Variant, p = 0.7, list = FALSE)
vino.train=data.pred[spl1,]
vino.test=data.pred[-spl1,]

#Fit the model with these important variables
mymodel<-rpart(as.factor(Variant)~., method="class", data = vino.train)
pred=predict(mymodel, data.pred)
mymodel.prob <- predict(mymodel, vino.test, type="prob")
threshold=1570/(1570+4461)
mymodel.pred = rep("red", nrow(vino.test))
mymodel.pred[which(mymodel.prob[,2] > threshold)] = "white"
confusionMatrix(factor(mymodel.pred),vino.test$Variant)
```

As we can see, the prediction accuracy is around 90%, which is a very high accuracy.